---
title: 'Craigs list used car post '
author: "Guoxin Li"
output: html_document
---



```{r setup}
library(knitr) # for changing workspace
library(stringr) # for using some regex functions
opts_chunk$set(root.dir = 'I:/R Data/141')
options(width = 110) 
load("vehicles.rda")
vehicle = vposts     # make a copy of the orignial data
body = vehicle$body
model_google = read.csv("model-google.csv")
```


```{r function}
value = function(regex,data)
{
  regex_fun = regex
  
  # It will store all the matched values in the posts return as a list
  result_fun= str_extract_all(data, ignore.case(regex_fun))
  
  # If there are multiple values in the list, get the first one out, if there is no value, give it NA
  value_fun = sapply(result_fun,function(x) {               
       if (length(x)>0) x[1]
       else NA
       })
  
  # how many posts contain the pattern 
  number = as.numeric(table(is.na(value_fun))[1])   
  
  # the posts that don't have that pattern 
  check = data[is.na(value_fun)]                             
  return(list(value_fun,number,check))                     
  
}
```


**Extract the price being asked for the vehicle from the body column, if it is present, and check if it agrees with the actual price in the price column.**

```{r}
# regular expression for price
price = "\\$?(\\d{0,3}[,.]?)*\\d{1,3}"           
price_value = value(price,body)                            

# how many posts have price in body
price_value[2]

# remove "$", "," or "." and then transform to numeric
price_value = as.numeric(gsub("\\$|,|.", "", price_value[[1]]))

# compare the price from "body" and from price column
table(price_value == vposts$price)                                
```

**Extract a Vehicle Identication Number (VIN) from the body, if it is present. We could use this to  both identify details of the car (year it was built, type and model of the car, safety features, body style, engine type, etc.) and also use it to get historical information about the particular car. Add the VIN, if available,to the data frame. How many postings include the VIN?**

```{r}
# regular expression for vin number
vin = "VIN[ :-]? ?\\d*\\w{1,7}\\d{1,3}\\w{1,5}\\d{4,8}"           
vin_num = value(vin,body)  

# how many posts have price in body
vin_num[2]
```

```{r}
# rough way to get the vin number
vin = "[A-HJ-NPR-Z0-9]{17}"                                     
vin_num1 = value(vin,body)  

# how many posts have price in body
vin_num1[2]

# look it the value 
 VIN = data.frame(vin_num1[[1]])
#View(VIN)
 vposts$vin = vin_num[[1]]
```

**Extract phone numbers from the body column, and again add these as a new column. How many posts include a phone number?**

```{r}
# regular expression for vin number
phone = "\\(?\\d{3}\\)?[ ]?[ -]?\\d{3}-? ?\\d{4}"               
phone_num = value(phone,body) 

# how many posts have price in body
phone_num[2]

```

**Extract email addresses from the body column, and again add these as a new column. How many posts include an email address?**

```{r}
# regular expression for vin number
email = "[[:alnum:]|[:punct:]]+@[[:alnum:]|[:punct:]]+?\\.(com|net|org|edu|gov){1}"     
email_add = value(email,body)

# how many posts have price in body
email_add[2]

# grep the posts have "email"
email = body[grepl("email", body, ignore.case = TRUE)]  

# how many posrs have email
length(email)

# look at some posts have "email"
email[1:3]
```

**Find the year in the description or body and compare it with the value in the year column.**

```{r}
# regular expression for year
year = " ?(19|20)\\d{2} ?"       
year_b = value(year,body)

# how many posrs have year
year_b[2]

# compare to the year column
year_num = as.numeric(year_b[[1]])
table(year_num == vposts$year) 
```

**Determine the model of the car, e.g., S60, Boxter, Cayman, 911, Jetta. This includes correcting mis-spelled or abbreviated model names. You may find the agrep() function useful. You should also use statistics, i.e., counts to see how often a word occurs in other posts and if such a spelling is reasonable, and whether this model name has been seen with that maker often. When doing these questions, you will very likely have to iterate by developing a regular expression, and seeing what results it gives you and adapting it. Furthermore, you will probably have to use two or more strategies when looing for a particular piece of information. This is expected; the data are not nice and regularly formatted.**

```{r}

# get the model from the "title" column
#title = "\\d{2,4} ([A-z]+[:punct:]?[A-z]+) ([A-z0-9]+)" 
#new_title  = value(title,vposts$title) 
#model = gsub(".*\\d{2,4} ([A-z]+) ([A-z0-9]+).*" , "\\2", new_title[[1]])
#model = casefold(model, upper = F)

# save it as csv (won't do it again, since I already saved it)
#vposts$model = model
#write.csv(vposts[,c(1,28)], file = "model.csv")

# read the sort data from google refine
#model_google = read.csv("model-google.csv")

# redefine model column
vposts$model = model_google[,3]

```

**Pick two models of cars, each for a different car maker, e.g., Toyota or Volvo. For each of these, separately explore the relationship between the price being asked for the vehicle, the number of miles (odometer), age of the car and condition. Does location (city) have an effect on this? Use a statistical model to be able to suggest the appropriate price for such a car given its age, mileage, and condition. You might consider a linear model, k-nearest neighbors, or a regression tree. You need to describe why the method you chose is appropriate? what assumptions are needed and how reasonable they are? and how well if performs and how you determined this? Would you use it if you were buying or selling this type of car?**

```{r}

# add age columns to the data
vposts$age = 2015- vposts$year

# Print out conditions so we can cut and paste them into smaller
conditions = levels(vposts$condition)
conditions = sprintf('"%s",\n', conditions)
cat(conditions)

# Define new categories. (code from homework1 solution for question 16)
new_cats = list(
  excellent = c("excellent"),
  good = c("good", "very good"),
  "like new" = c("like new", "mint", "new", "pre owned", "pre-owned", "preowned", "preownes"),
  used = c("0used", "used"),
  fair = c("fair", "nice", "nice teuck"),
  salvage = c("complete parts car, blown engine", "front side damage", "hit and run :( gently", 
              "muscle car restore", "needs bodywork", "needs restoration!", "needs restored",
              "needs total restore", "needs work", "needs work/for parts", "nice rolling restoration",
              "not running", "parts", "project", "project car", "rebuildable project", "restoration",
              "restoration project", "restore", "restored", "salvage", "rough but runs"),
  other = c("207,400", "ac/heater", "carfax guarantee!!", "certified", "honnda", "superb original" )
)

# Convert conditions to new categories.
vposts$new_cond = vposts$condition
levels(vposts$new_cond) = c(levels(vposts$new_cond), "other")

for (i in seq_along(new_cats)) {
  new_cat = names(new_cats)[[i]]
  vposts$new_cond[vposts$new_cond %in% new_cats[[i]]] = new_cat
}

vposts$new_cond = factor(vposts$new_cond)
```

```{r}
sort(table(vposts$model), decreasing = TRUE)[1:5]
```

```{r}

# get the  camry posts 
camry = vposts[vposts$model == "camry",]
camry = camry[!is.na(camry$model),]

# take out the NAs
remove_na = function(data)
{
  data = data[!is.na(data$odometer),]
  data = data[!is.na(data$price),]
  data = data[!is.na(data$new_cond),]
}

camry = remove_na(camry)

# plot the price with odometer, age and condition separately
plot_relation = function(data)
{
plot(data$odometer, data$price, ylab= "Price (US dollar) ", xlab = "Odometer (mile)", main = "Relation between Price and Odometer")
plot(data$age, data$price, ylab = "Price (US dollar) ", xlab = "Age (year)", main = "Relation between Price and Age")
plot(factor(data$new_cond), data$price, ylab= "Price (US dollar) ", xlab = "Condition", main = "Relation between Price and Condition")
}

par(mfrow=c(2,2))
plot_relation(camry)
```


```{r}
# take out the outliers
# see the 50 largest odometer
tail( sort(camry$odometer), 50 )

# I take odometer that no more than 30,000 and zero value
camry = camry[camry$odometer<300000 & camry$odometer != 0,]

#log trasformation
log_plot = function(data)
{
  plot(data$odometer, log(data$price), ylab= "Price (log(US dollar)) ", xlab = "Odometer (mile)", main = "Relation between log(Price) and Odometer")
  plot(data$age, log(data$price), ylab= "Price (log(US dollar)) ", xlab = "Age (year)", main = "Relation between log(Price) and Age")
}
par(mfrow=c(2,1))
log_plot(camry)
```


```{r}
# split price by city
pri_by_city = split(camry$price, camry$city)

# draw boxplot
par(mfrow=c(1,1))
boxplot(pri_by_city, col = "salmon")
title(" Camry Price Distribution by City", ylab = "Dollars" , xlab = "Cities")
```


```{r}
# fit a regression model
fit_model = lm(log(price)~odometer+age+factor(new_cond), data = camry)

# diagnostics for model
par(mfrow=c(2,2))
plot(fit_model)
```


```{r}
summary(fit_model)
anova(fit_model)
```

```{r}
# use cross validation to test the model
test_model = function(data)
{
  # split train and test data set
  index = sample(1:nrow(data), nrow(data)/5)
  train = data[-index,]
  test = data[index,]
  
  # fit the model
  fit_model = lm(log(price)~odometer+age+factor(new_cond), data = train)
  
  # format test data for prediction
  testdata = data.frame(with(test,cbind(odometer, age)))
  testdata$new_cond = test$new_cond
  
  # get the predict price
  pred_price = exp(predict(fit_model, testdata, type="response") )
  
  # plot the true and predict price
  plot(pred_price, type = "l", ylab = "Price (log(US dollar)) ", xlab = "Index", main = "True value versus predict value")
  points(test$price, col = "red", type = "l")
  legend("topleft", legend = c("True price", "Predict price"), col= c("black", "red"), pch=1, cex = 0.6)
  
}

# do it twice
par(mfrow=c(1,2))
test_model(camry)
test_model(camry)

```


```{r}
# get the  camry posts 
civic= vposts[vposts$model == "civic",]

# take out the NAs
civic = remove_na(civic)

# plot the price with odometer, age and condition separately
par(mfrow=c(2,2))
plot_relation(civic)
```

```{r}
# take out the outliers
# I take odometer that no more than 50,000
civic = civic[civic$odometer<500000 | civic$odometer == 0,]

#log trasformation
par(mfrow=c(1,2))
log_plot(civic)
```

```{r}
# split price by city
pri_by_city = split(civic$price, civic$city)

# draw boxplot
par(mfrow=c(1,1))
boxplot(pri_by_city, col = "salmon")
title(" Civic Price Distribution by City", ylab = "Dollars" , xlab = "Cities")
```

```{r}
# fit a regression model
fit_model = lm(log(price)~odometer+age+factor(new_cond), data = civic)

# diagnostics for model
par(mfrow=c(2,2))
plot(fit_model)
```


```{r}
# remove outlier
civic = civic[!rownames(civic) == "posted18193",]

# fit regression model agian
fit_model = lm(log(price)~odometer+age+factor(new_cond), data = civic)

summary(fit_model)
anova(fit_model)
```


```{r}
# use cross validation to test the model do it twice
par(mfrow=c(1,2))
test_model(civic)
test_model(civic)

```